Data Recovery and Pulse Position Modulation with a Photon Number Resolving SNSPD

### 2023-10-18 manuscript.tex
overleaf update

Deep Space Optical Communication has been a growing field of study in recent years, as researchers look for ways to communicate with spacecraft that are far away from Earth. The main challenge in deep space optical communication is the high loss and distance that the link must traverse. This limits the communication from the spacecraft, as it is limited by the power available on the spacecraft. This means that the communication protocol is limited by the number of bits that can be sent per unit of energy on the spacecraft, also known as photon information efficiency or bits per photon.

@@

The Deep Space Optical Communicaiton (DSOC) project managed by the Jet Propulsion Laboratory (JPL) aims to demonstrate optical communication using PPM with the Psyche spacecraft from distances of \textbf{\hl{XX to XX}} Au.

For larger M, more data may be sent with a single optical pulse, thereby allowing a spacecrat to that is severely power limited to send more data over a high loss optical channel back to earth. This is quantified through the photon information efficiency $c_p = C/E$ where $C$ is the link capacity and $E$ is the photon cost per optical pulse

The existing deep space optical communication project uses M up to $2^8 = 256$, meaning that 8 bits of data may be carried in each optical pulse. M values as high as 19 have been demonstrated in the lab\cite{essiambre2023record}. However, the number of time bins needed per optical pulse scales exponentially with the number of bits transmitted per pulse. Therefore, for a given fixed clock rate and time bin duration, the PPM data rate decreases dramatically for higher M values.

We focus on demonstrating moderately large PIE, while also increasing the clock rate of the sytem by an order of magnitude relative to the DSOC platform (2\textasciitilde GHz). By operating at both higher clock rate and PIE than DSOC, this system exemplifies how future iterations of DSOC may send data more quickly but also over greater distances with the same power budget.

The increase in clock rate is made possible due to recent advancements in Niobium Nitride SNSPDs\textasciitilde{}\cite{Colangelo2023}. Low jitter performance is achieved by incorporating impedance matching tapers for efficient RF coupling, resulting in higher slew rate pulses, and by enabling RF pulse readout from both ends of the nanowire. The dual-ended readout allows for the cancellation of jitter caused by the variable location of photon arrival along the meander when the differential signals are recombined with a balun.
These detectors acheive jitter as low as 50 ps at the FW(1/100)M level, making them suitable for the demonstration of PPM with 50 ps slot widths and a 20 GHz clock.

@@

The work here highlights the application of low jitter single-photon detectors for optical communication, which is impactful for deep-space optical communication as well as classical communication in quantum networks. Although single-photon counting is well estanblished for deep-space optical communication\textasciitilde{}\cite{Laser lunar, DSOC} so far it has not been ulitized in quantum networks, mainly due to the use of SFP modules and DWDMs. However, with the eachievement of high data rates recently achieved with photon-counting classical communication, these approaches can now be seriously considered for quantum networks. The main driver is would be the deduction of optical power in neighbouring DWDM channels, which ultimately lowers the Raman scattered photons into the quantum channel \cite{EraerdsRaman}
\textbf{\hl{Calculate reduction in power from state of the art SFP modules}}

@@

>>>>

Deep space optical communication has been a growing field of study in recent years, as researchers look for ways to communicate with spacecraft that are far away from Earth. A key challenge in the development of this technology is closing the communication link over extremely large distances and high loss. This must be done given a restricted power budget available on the spacecraft, and therefore requires the use of highly efficient optical communication protocols that optimize the number of bits sent per unit of energy.

@@

The Deep Space Optical Communicaiton (DSOC) project managed by the Jet Propulsion Laboratory (JPL) aims to demonstrate optical communication using PPM with the Psyche spacecraft from distances of up to \textbf{\hl{XX}} Au.

For larger M, more data may be sent with a single optical pulse, thereby allowing a power limited spacecraft to send more data over a high loss optical channel back to earth. This is quantified through the photon information efficiency $c_p = C/E$ where $C$ is the link capacity and $E$ is the photon cost per optical pulse

The existing DSOC project uses M up to $2^8 = 256$, meaning 8 bits of data may be carried in each optical pulse. M values as high as 19 have been demonstrated in the lab~\cite{essiambre2023record}, but the number of time bins needed per frame scales exponentially with the number of bits transmitted per pulse. Therefore, for a given fixed clock rate and time bin duration, the PPM data rate decreases dramatically for higher M values.

We demonstrate a high clock rate PPM protocol in the lab based on modulating a mode-locked laser and receiving pulses with a low jitter superconducting nanowire single photon detector (SNSPD) (fig.~\ref{fig:intro} (a)).We focus on demonstrating moderately large PIE, while also increasing the clock rate of the sytem by an order of magnitude relative to the DSOC platform (2\textasciitilde GHz). By operating at both higher clock rate and PIE than DSOC, this system exemplifies how future iterations of DSOC may send data more quickly but also over greater distances with the same power budget.

The rate increase is possible due to recent advancements in Niobium Nitride SNSPDs~\cite{Colangelo2023}. These acheive low jitter performance by incorporating impedance matching tapers for efficient RF coupling, resulting in higher slew rate pulses, and by enabling RF pulse readout from both ends of the nanowire. The dual-ended readout allows for the cancellation of jitter caused by the variable location of photon arrival along the meander when the differential signals are recombined with a balun. These detectors achieve jitter as low as 50 ps at the FW(1/100)M level, making them suitable for the demonstration of PPM with 50 ps slot widths and a 20 GHz clock.

@@

The work here highlights the application of low jitter single-photon detectors for optical communication, which is impactful for deep-space optical communication as well as classical communication in quantum networks. Although single-photon counting is well estanblished for deep-space optical communication~\cite{Laser lunar, DSOC} so far it has not been ulitized in quantum networks, mainly due to the use of SFP modules and DWDMs. However, with the eachievement of high data rates recently achieved with photon-counting classical communication, these approaches can now be seriously considered for quantum networks. The main driver is would be the deduction of optical power in neighbouring DWDM channels, which ultimately lowers the Raman scattered photons into the quantum channel \cite{EraerdsRaman}
\textbf{\hl{Calculate reduction in power from state of the art SFP modules}}

@@

{}
